---
title: "Loops and Functions"
author: "Ruth Enriquez"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Part 1 For loops revisited

## Task 1

Create two sequences, one called weekdays that contains days of the week ("Monday", "Tuesday", "Wednesday", etc.) and one called `transects` that contains the series of transect names "Transect A", "Transect B,"Transect C". Write a nested for loop that creates a matrix containing the following

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friyay", "Saturday", "Sunday")
transects <- c("Transect A", "Transect B", "Transect C")

weekday_transect <- matrix(nrow = length(weekdays), ncol = length(transects))

for (i in seq_along(weekdays)) {
  for (j in seq_along(transects)) {
    weekday_by_transect <- paste0(weekdays[i],"-", transects[j])
    weekday_transect[i, j] <-weekday_by_transect
  }
}

weekday_transect
```

# Part 2 Functions introduced

## Task 2

Write a function called `force` that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2ms2 (recall: F=maF=ma), and returns a statement "The resulting force is \_\_\_ Newtons."

```{r}
force <- function(m, a) {
  paste("The resulting force is", m * a, "Newtons")
}

force(m=10, a=5)
```

## Task 3

The length:weight relationship for fish is: W=aLb, where L is total fish length (centimeters), W is the expected fish weight (grams), and a and b are species-dependent parameter values (shown below for several fish from Peyton et al. 2016).

Recreate the table above as a data frame stored as `fish_parms`

```{r}
fish_parms <- tribble(
  ~sci_name, ~common_name, ~a_est, ~b_est,
  "Chanos chanos", "Milkfish", 0.0905, 2.52,
  "Sphyraena barracuda", "Great barracuda", 0.0181, 3.27,
  "Caranx ignobilis", "Giant trevally", 0.0353, 3.05
)

fish_length <- seq(from = 0, to = 100, by = 1)
```

Then, write a function called `fish_weight` that allows a user to only enter the *common name* (argument `fish_name`) and *total length* (argument `tot_length`) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and length.

```{r}
fish_weight <- function(name, tot_length) {
  fish_subset <- filter(fish_parms, common_name == name)
  print(fish_subset$a_est*tot_length^fish_subset$b_est)
}

fish_weight(name = "Milkfish", 10)

milkfish_weights <- fish_weight(name = "Milkfish", tot_length = fish_length)

print(milkfish_weights)
```

## Task 4

```{r}
wave_power <- function(H, Period) {
  0.5*H^2*Period
}
wave_height <- seq(from = 0, to = 3, by = 0.2)

wave_power(wave_height, 8)
```
